R for Clinical Study Reports and Submission

### 2023-07-16 tlf-disposition.qmd
update `tlf-population.qmd`

To prepare a simplified disposition of participants table using the `adsl` dataset, the following variables are utilized:

@@

By using `tbl_disp` as input, we create generate the disposition table using following code.
The `r2rtf` elements that were not covered in the previous sections are highlighted.
Here, the `rtf_title` function is utilized to define the title of the table. 
For the `title` argument, 
you can provide a vector where each value represents a separate line.

>>>>

To prepare a simplified disposition of participants table using the `adsl` dataset, the following variables are used:

@@

By using `tbl_disp` as input, we can generate the disposition table using following code.
The `r2rtf` elements that were not covered in the previous sections are highlighted.
Here, the `rtf_title` function is utilized to define the title of the table. 
For the `title` argument, 
you can provide a vector where each value represents a separate line.

### 2023-07-16 tlf-population.qmd
update `tlf-population.qmd`

In this chapter, we illustrate how to create a summary table for
the analysis population for a study.

@@

The first step is to read relevant datasets into R. For the analysis
population table, all the required information is saved in the ADSL
dataset. We can use the `haven` package to read the dataset.

@@

We illustrate how to prepare a report data for a simplified analysis
population table using variables below:

@@

Before we write the analysis code, let's discuss the possibility of reusing R
code by writing helper functions.

As discussed in [R for data
science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function whenever you've copied and
pasted a block of code more than twice".

In @sec-disposition, there are a few repeating steps to:

- Format the percentages using the `formatC()` function.
- Calculate the numbers and percentages by treatment arm.

@@

To format numbers and percentages, we create a function called
`fmt_num()`. It is a very simple function wrapping `formatC()`.

@@

The main reason to create the `fmt_num()` function is to enhance the
readability of the analysis code.

For example, we can compare the two versions of code to format the
percentage used in @sec-disposition and `fmt_num()`.

@@

The function can be enhanced in multiple ways, but here we only focus on
simplicity and readability. More details about writing R functions can
be found in the [STAT 545
course](https://stat545.com/functions-part1.html).

@@

By using the `count_by()` function, we can simplify the analysis code as
below.

@@

## Analysis code

With the helper function `count_by`, we can easily prepare a report
dataset as

@@

Now we combine individual rows into one table for reporting
purpose. `tbl_pop` is used as input for r2rtf to create the final
report.

@@

We define the format of the output using code below.

@@

The procedure to generate an analysis population table can be summarized
as follows:

- Step 1: Read data (i.e., `adsl`) into R.
- Step 2: Bind the counts/percentages of the ITT population, the
  efficacy population, and the safety population by row using the
  `count_by()` function.
- Step 3: Format the output from Step 2 using r2rtf.

>>>>

This chapter provides a step-by-step guide on creating a simple table for
the analysis population of a study in RTF format.
The expected output of the table is as follows:

@@

## Input data

The initial step involves reading the ADSL datasets into R.

@@

the following variables are utilized:

@@

Before we writing the analysis code, 
let's explore the possibility of maximizing 
code reuse in R by creating helper functions.

::: {.callout-note}
As emphasized in the book [R for data
science](https://r4ds.had.co.nz/functions.html#when-should-you-write-a-function),
"You should consider writing a function whenever you've copied and
pasted a block of code more than twice".
:::

In @sec-disposition, there are a few repetitive steps involved:

- Formatting the percentages using the `formatC()` function.
- Calculating  the numbers and percentages by treatment arm.

@@

To address this, we introduce two ad-hoc functions that 
will assist us in creating tables throughout the remainder of this book.

The first function, named `fmt_num()`, is a straightforward implementation that 
wraps `formatC()` to format numbers and percentages.

@@

One purpose of creating the `fmt_num()` function 
is to improve the readability of the analysis code.

To illustrate this, let's compare two versions of code for formatting percentages: 
the one used in @sec-disposition and the alternative approach using fmt_num().

@@

To compute the numbers and percentages of participants in different groups,
we offer a straightforward (albeit not robust) 
wrapper function called `count_by()`. 
This function depends on the dplyr and tidyr packages.

::: {.callout-note}
While there are several ways to improve the function, 
our current focus is on enhancing simplicity and readability. 
A good tutorial can be found from the  [STAT 545
course](https://stat545.com/functions-part1.html).
:::

@@

The `count_by()` function help us simplify the analysis code:

@@

## Output data

The objective of this section is to create a dataset suitable for generating the expected report. 
This process will consist of several steps to generate each component of the reporting dataset.

With the aid of the helper function `count_by()`, 
preparing a report dataset becomes a straightforward task.

@@

@@

## Reporting

By using `tbl_pop` as input, we can generate the analysis population table using following code.

@@
